---
# this is the Rmarkdown header. Use the YAML syntax. ---YAML block ---.
title: "Predicting Employee Performance" 
subtitle: "Groupwork Use Case 4 - People Analytics"
author: "Substitute Group_X by your Group number in the name of the .Rmd file"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide

# the following lines tell R to store the output in the subdirectory reports
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
---

# Intro

In this workshop, you will explore multiple models to predict individual, team and organizational performance. You will apply multiple regression, stepwise regression, and dummy variable techniques to performance data from various sectors. Your goal is to derive actionable insights about what drives performance and how organizations can use these insights ethically and strategically.

# Setup

Within the project directory, create the following folders if they do not already exist:

-   **Databases**: to store all data sources provided in Moodle (Excel files).\
-   **data**: to store any ad-hoc data created during the analysis. The native R data format is .rds.\
-   **report**: to store the HTML output file generated when knitting the R Markdown file.

```{r setup, include=FALSE}
install.packages("tidyverse")
install.packages("readxl")
install.packages("lm.beta")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
```

# Prompt 1: Predicting Customer Loyalty (Financial Services)

### Tasks

-   Import 'customer_satisfaction.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate a multiple regression model with Custloyalty as the dependent variable. Predictors: Sat1–Sat4 and SexOfSalesperson.\
-   Interpret R², F-statistic and beta coefficients.\
-   Recommend areas for sales improvement.\

### Deliverables

-   Model output and standardized coefficients.\
-   Strategic recommendation for salesperson development

## Solution

### Model output and standardized coefficients

```{r prompt1.1}
library(readxl)
library(lm.beta)

# Carga de datos
df1 <- read_excel("Databases/customer_satisfaction.xlsx")

# Modelo
model1 <- lm(Custloyalty ~ Sat1 + Sat2 + Sat3 + Sat4 + SexOfSalesperson, data = df1)
summary(model1)

# Coeficientes estandarizados
lm.beta(model1)
```

The R\^2 is 0.3257, which means that approximately 32.6% of the variance in customer loyalty is explained by this model. While there are other unknown factors influencing loyalty, capturing 33% of the variance with just 4 satisfaction metrics and the sex of the salesperson is meaningful in business.

To continue, we can see that the F-statistic is 176.7 with a p-value\<2.2e-6. This means that the overall regression model is highly significant as a model.

If we look at the beta coefficients we can see that Sat1 is 0.306, with a p-value\<2e-16 and Sat3 is 0.290, with a p-value\<2.2e-16. They are the most significant independent variables of the model. On the other hand, Sat2 is -0.0295 and Sat4 is 0.0475, which are not statistically significant at the 5% level, meaning changes in these areas do not clearly impact loyalty. Lastly, SexOfSalesperson is significant (p-value = 0.025), but it has a very weak effect size (beta= -0.043) compared to the significant satisfaction variables.

### Strategic recommendation for salesperson development

Based on the model, we observe that having a salesperson who understands the customer's needs and is friendly are the variables that most positively affect customer loyalty. In other words, the more friendly and able to understand customer needs the employees are, the more the company will be able to grow and retain its customer base.

On the other hand, whether the salesperson seems very confident or whether they are highly knowledgeable does not have a statistically significant impact on loyalty. While seeming overly confident even showed a slight negative trend, it is not a defining factor. Therefore, we recommend that the company should not focus resources on developing pure technical knowledge, but rather prioritize active listening, empathy and a welcoming attitude.

Finally, the model indicates that customer loyalty is slightly higher when the salesperson is female rather than male. To leverage this insight, the company could implement a mentoring program where top-performing salespeople share their best practices so that the entire sales force can learn from the company's most successful employees.

# Prompt 2: Predicting Customer Reinvestment (Financial Services)

### Tasks

-   Import 'customer_satisfaction.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Replicate Prompt 1 using InvestMore as the dependent variable.\
-   Compare results and R².\
-   Discuss implications for upselling and cross-selling strategy.\

### Deliverables

-   Model summary and interpretation.\
-   Recommendations on customer retention strategies.\

## Solution

### Model summary and interpretation

```{r prompt2.1}
model2 <- lm(INvestMore ~ Sat1 + Sat2 + Sat3 + Sat4 + SexOfSalesperson, data = df1)
summary(model2)

```

### Statistical Analysis & Interpretation

-   $R^2$ value is **0.1704**. This indicates that only **17%** of the variance in a customer's intent to "Invest More" is explained by the model.

-   **Comparison with Prompt 1:** This is significantly lower than the $R^2$ for Customer Loyalty (32.6%). This suggests that while satisfaction with a salesperson influences **loyalty** (emotional connection), the decision to **invest more money** is likely driven by external factors not in this model, such as interest rates, personal income, or product performance.

-   **F-statistic:** 72.53 ($p < 2.2e-16$). Despite the lower $R^2$, the model is still statistically significant, meaning the predictors as a group do have a reliable relationship with the intent to reinvest.

### **Significant Drivers (p \< 0.05)**

-   **Sat1:** Understanding Your Needs (The "Consultant" Driver)

    -   **Statistical Meaning:** With a coefficient of **0.175**, this is the strongest predictor of increased investment.

        This measures **active listening and empathy**. Customers are most likely to increase their financial commitment when they feel the salesperson isn't just selling a product, but is accurately diagnosing their specific financial situation.

-   **Sat3: Having a Recommendation (The "Guidance" Driver)**

    -   **Statistical Meaning:** This is also highly significant (Coefficient **0.139**).

        This measures **decisiveness and value-add**. It’s not enough to just understand the customer (Sat1); the salesperson must provide a clear "call to action" or a specific path forward. Customers invest more when they are given a clear direction rather than a list of overwhelming options.

### **Non-Significant Drivers (p \> 0.05)**

-   **Sat2: Confidence (The "Persona" Driver)**

    -   **Statistical Meaning:** No significant impact on reinvestment.

        Appearing confident does not translate into more capital from the client. This suggests that "surface-level" sales charisma is less important than the actual substance of the advice given.

-   **Sat4: Knowledgeable (The "Expertise" Driver)**

    -   **Statistical Meaning:** No significant impact.

        This is a surprising but common finding in Financial Services. While a salesperson must be competent, "being knowledgeable" is often viewed as a **baseline requirement**. High knowledge alone doesn't inspire a customer to double their investment if the salesperson fails to apply that knowledge to the customer's specific needs (Sat1).

-   **SexOfSalesperson:**

    -   **Statistical Meaning:** Not significant ($p = 0.651$).

        The gender of the salesperson has no bearing on the volume of capital a customer decides to reinvest.

### Recommendations on customer retention strategies

Based on the regression analysis of customer behavior, the organization’s strategy must transition from providing **Information** to delivering **Actionable Insight**. While the current sales force is perceived as knowledgeable, technical expertise alone is not driving capital growth. To move customers from "uncertainty" to "doubling investment," the following four strategies are recommended:

#### 1. Re-Focusing Sales Training:

The data indicates that **Knowledge (Sat4)** is not a significant predictor of increased investment, whereas **Understanding Needs (Sat1)** is a primary driver. This suggests that customers view technical knowledge as a baseline requirement, not a reason to increase their financial commitment.

-   **Strategy:** HR should pivot training away from memorization of product manuals.

-   **Action:** Start workshops focused on **Active Listening**. We want our team to spend more time asking smart questions and less time reading from a script.

#### 2. Implementation of a "Structured Recommendation" Protocol

**Sat3 (Having a Recommendation)** emerged as a statistically significant driver of reinvestment ($p < 0.001$). Customers are more likely to commit more capital when they are provided with a clear path forward rather than a menu of options.

-   **Strategy:** Empower the sales force to move beyond "informed service" into "bold advisory."

<!-- -->

-   **Action:** Establish a **"Next Best Action"** protocol. Every customer interaction must conclude with a specific, tailored recommendation. This way, by closing the gap between "what we offer" and "what you should do," the firm can capture the "Not Sure" segment of the customer base (CustomerLoyalty).

#### 3. Leveraging Loyalty as a Lead Indicator for Wallet-Share

The model for **Customer Loyalty (**$R^2 = 0.33$) is substantially stronger than the model for **Increased Investment (**$R^2 = 0.17$). This suggests that while salespeople have a high degree of influence over whether a customer stays with the firm, external market factors play a larger role in how much they invest. It is much easier for our team to make a customer **loyal** than it is to convince them to **double their money**. We should use "Friendliness" and "Understanding" to keep them in the door first

-   **Strategy:** Build the relationship first. Once the customer trusts us (Loyalty), they are much more open to our recommendations for bigger investments.

-   **Action:** Focus retention efforts on **Empathy and Needs-Matching** to secure loyalty first. Once a high "Loyalty Score" is achieved, trigger an automated or personal outreach focused on the "Recommendation" (Sat3) to bridge the gap from retention to upselling.

#### 4. Evidence-Based Recruitment and Persona Optimization

The analysis proves that **Confidence (Sat2)** and **Gender (SexOfSalesperson)** are statistically irrelevant to the customer's decision to invest more. The "traditional" high-confidence, high-pressure sales persona does not correlate with actual business growth in this sector.

-   **Strategy:** Broaden the recruitment profile to prioritize behavioral competencies over sales stereotypes.

-   **Action:** Adjust hiring criteria to look for people who are naturally good at empathy and analytical thinking. The ideal candidate is not necessarily the most "confident" in the room, but the one most capable of synthesizing client data into a coherent, persuasive recommendation.

# Prompt 3: Linking Employee Engagement to Customer Loyalty

### Tasks

-   Import 'aggregate_sales_customer_loyalty.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Predict average customer loyalty per salesperson. Predictors: qEng1, qEng2, qMan1, Talent2014, Sex, Age, Manager.\
-   Evaluate significance and limitations.\
-   Discuss engagement-performance links.

### Deliverables

-   Regression output and interpretation.\
-   Organizational insight and caveats.\

## Solution

### Regression output and interpretation

```{r prompt3.1}


```

### Organizational insight and caveats

# Prompt 4: Predicting Individual Performance Ratings (Manufacturing)

### Tasks

-   Import 'perf_2014_to_2015_sick_2014.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Predict 2015 performance using 2014 predictors. Variables: POS2014, DistJust2014, ProcJust2014, JOBSAT2014, ValueFit2014, JobStrain2014, SickDays2014.\
-   Interpret R² and beta values.\
-   Discuss what drives performance improvements.

### Deliverables

-   Model summary and insights.\
-   HR recommendations based on findings.\

## Solution

### Model summary and insights

```{r prompt4.1}
# Importar datos [cite: 31]
df_4 <- read_excel("Databases/perf_2014_to_2015_sick_2014.xlsx")

# Modelo con predictores de actitud y salud [cite: 33]
model4 <- lm(PerformanceRating2015 ~ POS2014 + DistJust2014 + ProcJust2014 + JOBSAT2014 + 
               ValueFit2014 + JobStrain2014 + SickDays2014, data = df_4)

summary(model4)

```

### HR recommendations based on findings

# Prompt 5: Stepwise Regression for Performance Prediction

### Tasks

-   Import 'perf_2014_to_2015_sick_2014.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Use stepwise regression to select significant predictors.\
-   Compare model fit with Prompt 4.\
-   Evaluate parsimony and implications for policy.

### Deliverables

-   Stepwise model output.\
-   Strategic use of simpler models in HR.\

## Solution

### Stepwise model output

```{r prompt5.1}
library(MASS)

model5_step <- stepAIC(model4, direction = "both")
summary(model5_step)
```

### Strategic use of simpler models in HR

When comparing this stepwise model to the previous model, we observe the principle of model efficiency. The stepwise regression algorithm automatically removed four non-significant predictors (JOBSAT, ValueFit, DistJust and ProcJust), leaving only three key variables: POS2014, JobStrain2014, and SickDays2014. Despite having fewer variables, the adjusted R\^2 slightly increased from 0.0675 to 0.0692. This means our simpler model is mathematically more efficient, explaining the variance in performance just as well by cutting out the statistical noise.

This has an impact on HR strategies. If HR only needs to measure three variables instead of seven to predict performance, they can deploy much shorter employee surveys. This improves response rates and saves employees time.

In addition, a simpler model provides management with actionable, targeted goals. Instead of trying to improve every aspect of the employee experience at once, HR leadership can channel their budget into just two main areas: implementing programs that increase Perceived Organizational Support and redesigning roles to reduce Job Strain.

Lastly, it is much easier to explain a simple, three-variable model to non-technical stakeholders and business leaders than a complex model with multiple conflicting or insignificant metrics.

# Prompt 6: Controlling for Prior Performance Ratings

### Tasks

-   Import 'perf_2014_to_2015_sick_2014.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Include 2014 performance rating as a predictor.\
-   Identify whether other variables still predict 2015 performance.\
-   Explain improvement prediction logic.

### Deliverables

-   Output with adjusted R² and key drivers.\
-   Change-focused HR insight.\

## Solution

### Output with adjusted R² and key drivers

```{r prompt6.1}

model6 <- lm(PerformanceRating2015 ~ PerformanceRating2014 + POS2014 + DistJust2014 + ProcJust2014 + 
               JOBSAT2014 + ValueFit2014 + JobStrain2014 + SickDays2014, data = df_4)
summary(model6)
```

-   **Model Fit:** The Adjusted $R^2$ is **0.2027**. This is a significant improvement from the model in Prompt 4 (which was only \~0.06). By including 2014 performance, we now explain about **20.3%** of the variance in 2015 performance.

-   **The Power of Stability:** **PerformanceRating2014** is the strongest predictor ($Estimate = 0.370$, $p < 0.001$). This confirms that the best predictor of future performance is past performance.

-   **Key Drivers of "Change":** Even after controlling for how well someone performed last year, two variables remain significant:

    1.  **JobStrain2014 (**$p = 0.0107$): Interestingly, job strain has a positive coefficient (0.171). This suggests that in this manufacturing context, higher pressure is associated with higher performance ratings, possibly indicating a "high-performance/high-stress" culture.

    2.  **SickDays2014 (**$p = 0.0357$): This has a negative impact. For every extra sick day taken in 2014, 2015 performance is expected to drop by 0.016 units, even if the employee was a high performer previously.

### Change-focused HR insight

#### 1. Understanding "Performance Momentum"

In HR analytics, adding a prior-year baseline changes the question from *"Who is a high performer?"* to **"What causes an employee's performance to improve or decline?"** Because past performance is so dominant, HR should focus less on variables that became non-significant (like Job Satisfaction or Value Fit) when trying to predict *short-term* performance changes. Those factors might matter for long-term retention, but they aren't the primary drivers of year-over-year performance swings in this dataset.

#### 2. The Health-Performance Lingering Effect

The significance of **SickDays2014** is a critical warning sign. It shows that health issues have a "carry-over" effect. An employee's illness last year predicts a decline in performance this year, regardless of their past talent level.

-   **Strategy:** HR should view sick leave not just as a temporary absence, but as a lead indicator of future performance risk. "Well-being" programs should be framed as "Performance Protection" investments.

#### 3. The Risk of the "Strain" Culture

The model shows that **Job Strain** is linked to higher performance. While this might look "good" on a spreadsheet (higher stress = higher output), it is a dangerous management strategy.

-   **Caveat:** If high performance is only achievable through high strain, the organization is likely heading toward burnout. HR must investigate if these high performers are also the ones taking more sick days. If so, the "gain" in performance from strain is being cancelled out by the "loss" in performance from health issues.

#### 4. Precision Coaching

By using this model, HR can identify "Outliers"—employees whose 2015 performance was much higher or lower than their 2014 baseline would suggest.

-   **Action:** Instead of universal training, HR can target interventions. For example, if an employee’s performance dropped despite low sick days and low strain, managers can look into other "non-model" factors like a change in leadership or personal issues.

# Prompt 7: Predicting Sickness Absence

### Tasks

-   Import 'perf_2014_to_2015_sick_2014.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Switch dependent variable to SickDays2014.\
-   Use attitude variables as predictors.\
-   Analyze non-significant results and discuss potential.

### Deliverables

-   Regression output.\
-   Interpretation and potential data limitations.\

## Solution

### Regression output

```{r prompt7.1}


```

### Interpretation and potential data limitations

# Prompt 8: Exploring Performance Patterns by Demographics

### Tasks

-   Import 'MI_data_perf.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Predict performance rating using SalaryBenchmark, Gender, AgeCat, TenureCat and Country dummies.\
-   Highlight impact of gender and country.\
-   Identify signs of bias in appraisal systems.

### Deliverables

-   Regression output.\
-   Recommendations to improve fairness in performance evaluation.\

## Solution

### Regression output

```{r prompt8.1}


```

### Recommendations to improve fairness in performance evaluation

# Prompt 9: Predicting Checkout Performance by Demographics

### Tasks

-   Import 'supermarket_checkout.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Predict ScanPminute using Gender, education, tenure, store, and age.\
-   Run model with age dummies to improve fit.\
-   Discuss age-related productivity insights.\
-   Address ethical concerns in applying the findings.

### Deliverables

-   Two models (with and without dummies).\
-   Strategic and ethical recommendations for workforce planning.\

## Solution

### Two models (with and without dummies)

```{r prompt9.1}
library(readxl)

# Importar los datos del supermercado
df_checkout <- read_excel("Databases/supermarket_checkout.xlsx")

# MODELO 1: Sin dummies (Predecir ScanPminute usando la edad como tendencia lineal)
model9_sin_dummies <- lm(ScanPminute ~ Gender + educ + Weeks + Store + AgeVariable, data = df_checkout)

# MODELO 2: Con dummies de edad (Usando las columnas que ya vienen en el Excel)
# Dejamos fuera 'age1' para que sirva como nuestra categoría de referencia base
model9_con_dummies <- lm(ScanPminute ~ Gender + educ + Weeks + Store + age2 + age3 + age4 + age5 + age6 + age7 + age8 + age9 + age10, data = df_checkout)

# Resultados
summary(model9_sin_dummies)
summary(model9_con_dummies)
```

### Strategic and ethical recommendations for workforce planning

When comparing both models, it is clear that Model 2 is significantly better. Its adjusted R\^2 is 0.3832 compared to 0.2485 in Model 1, demonstrating that the age variable does not have a linear relationship with productivity. By adding age brackets, we see that while a linear age variable was not significant, specific age categories are highly significant. Productivity peaks mid-career, specifically for employees in the age4 and age6, who show the highest increases in items scanned per minute (+7.944 and +7.313) compared to the youngest cohort. Interestingly, the oldest cohort (age10) is not significant, meaning their scanning speed is statistically comparable to the youngest entry-level cashiers.

Regarding other variables, weeks positively impacts productivity, while higher education levels slightly decrease scanning speed (which could indicate overqualification or lower job engagement). Finally, gender is a significant factor, with women scanning approximately 2.76 more items per minute than men.

-   Ethical recommendations: management must not use these findings to filter out male candidates, exclusively hire middle-age workers, or segregate demographics into different supermarket roles. Using demographics to restrict hiring or influence job placement violates equal opportunity employment laws and introduces systemic bias. Performance should always be evaluated on an individual basis, not predicted by demographic group membership.

-   Strategic recommendations:

    -   Instead of restricting hiring, the company should identify the specific techniques that make the middle-aged and female faster. By identifying these efficient behaviors, the company can standardize these methods and integrate them into the core onboarding program for all new hires.

    -   The performance gap could be related to the physical design of the checkout counter. HR and Operations should investigate if the current scanner height or bagging area reach inadvertently favors certain physical profiles. Improving workstation ergonomics could naturally boost the speed of all cashiers and reduce fatigue.

    -   To manage the store effectively, management should use performance data to balance shifts. Pairing high-efficiency mid-career cashiers with developing staff during peak hours ensures that customer wait times remain low, bottlenecks are avoided, and slower employees have opportunities for on-the-job observational learning.
