---
title: "Advanced Turnover Modelling and Business Case Development" 
subtitle: "Groupwork Use Case 3 - People Analytics"
author: "Ferrán Extremera, Irene Morales, Amalia Varo y Martina Iscar"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
editor_options: 
  markdown: 
    wrap: sentence
---

# Intro

Employee turnover is a critical concern for organizations due to its substantial financial and human costs.
HR functions frequently report turnover statistics, but descriptive reporting alone may lead to misdiagnosis of problems and poorly targeted interventions.
In this use case, you will apply inferential and predictive analytics to understand voluntary employee turnover at both individual and team levels in a multinational organization

# Setup

Within the project directory, create the following folders if they do not already exist:

-   **Databases**: to store all data sources provided in Moodle (Excel files).\
-   **data**: to store any ad-hoc data created during the analysis. The native R data format is .rds.\
-   **report**: to store the HTML output file generated when knitting the R Markdown file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(readxl)
```

# Prompt 1: Estimating differences in turnover by country

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Produce a cross-tabulation of Country × Leaver Status.\
-   Conduct a chi-square test of independence to assess whether turnover differs significantly by country. Interpret the results, paying attention to: expected vs observed frequencies; statistical significance; limitations related to small subgroup sizes.\

### Deliverables

-   Cross-tabulation table with expected frequencies.\
-   Chi-square test output.\
-   Interpretation: Should HR be concerned about country differences in individual turnover based on this analysis alone?

## Solution

```{r}
df1 <- read_excel("Databases/Individual_Turnover.xlsx")
head(df1)
```

```{r}
str(df1)
```

```{r}
summary(df1)
```

We need to treat the Gender and LeaverStatus variables carefully, as they have 2 (Gender) and 1(LeverStatus) NaN values.
Therefore, we must remove them primarily as a precaution.
Additionally, since it involves at most three rows of the dataset compared to the 1653 total rows, it is just a minimal deletion that aims to have a purer dataset and anticipate possible problems that these null values may cause in the models used throughout the practice.

```{r}
data <- df1 %>%
  drop_na(Gender, LeaverStatus)
```

```{r}
summary(data)
```

We verify that the null rows have been removed.
As a result we have a total of 1650 records for our analysis.

```{r}
str(data)
```

### Cross-tabulation table with expected frequencies

```{r prompt1.1}
tabla_cruzada <- table(
  "País" = data$Country, 
  "Estado_Rotación (0=Stay, 1=Leave)" = data$LeaverStatus
)

# Añadir una fila y columna de "Totales"
tabla_con_totales <- addmargins(tabla_cruzada)

# Imprimir en consola para que la veas tú mientras programas
cat("\n--- FRECUENCIAS OBSERVADAS ---\n")
print(tabla_con_totales)

# Crear la versión "bonita" para el reporte HTML final
knitr::kable(tabla_con_totales, 
             caption = "Employee Turnover Frequency by Country")
# Recordar cuando presentar poner el pais correspondiente al lado de número 
```

### Chi-square test output

```{r prompt1.2}
# 1. Asegurarnos de tener el test calculado
test_chi <- chisq.test(tabla_cruzada)
```

```{r}
cat("\n--- RESULTADO DEL TEST CHI-CUADRADO ---\n")
print(test_chi)
```

```{r}
print(test_chi$expected)
```

```{r}
# 1. Extraer los valores numéricos temporalmente}
stay_obv <- test_chi$observed[, 1]
leave_obv <- test_chi$observed[, 2]
stay_exp <- round(test_chi$expected[, 1], 1)
leave_exp <- round(test_chi$expected[, 2], 1)

# 2. Calcular la diferencia antes de convertir los números a texto
dif_stay <- stay_obv - stay_exp

# 3. Crear la tabla final uniendo Observado (Esperado)
tabla_completa <- data.frame(
  Country = rownames(test_chi$observed),
  
  # Usamos paste0 para unir el número real, un espacio, el paréntesis y el esperado
  `Stay_Obv_Exp` = paste0(stay_obv, " (", stay_exp, ")"),
  `Leave_Obv_Exp` = paste0(leave_obv, " (", leave_exp, ")"),
  
  Dif_Stay = dif_stay
)

# 4. Renombrar las columnas para que queden perfectas en el HTML final
# (R no deja poner paréntesis en el data.frame directamente de forma fácil)
colnames(tabla_completa) <- c("Country", "Stay: Obv (Exp)", "Leave: Obv (Exp)", "Dif_Stay")

# 5. Mostrar la tabla con un diseño elegante
# Añadimos align = "c" para que los números queden centrados y más legibles
knitr::kable(tabla_completa, 
             row.names = FALSE, 
             align = "c", 
             caption = "Complete Comparison: Retention and Turnover (Observed vs Expected) by Country")

```

**Test Validity and Limitations (The Warning)** When running the Chi-square test, R triggers a warning indicating that the approximation may be incorrect.
This happens due to an issue with small subgroup sizes.
For the Chi-square mathematical calculation to be highly reliable, a strict rule applies: the *expected* frequencies in every cell must be greater than 5.
As seen in our expected table, countries like Country 2 (4.56) and Country 4 (4.68) fall below this threshold, causing the statistic to lose reliability.

If ignored, this limitation could trigger false alarms for Human Resources, mistaking normal random variations in small offices for genuine retention issues.

To correct this statistically, we could use the **Fisher Exact Test**, which is designed for small subgroups.
However, because we have a considerable dataset (N=1650 records), computing all possible combinations requires massive memory and might cause an error.
A more viable statistical alternative is to run a **Monte Carlo Simulation**, which bypasses the strict formula by generating thousands of random scenarios to calculate a highly accurate $p$-value.

**The HR Perspective on the Warning:** From a pure Human Resources standpoint, the best practice is simply to group the data.
For instance, combining the smallest delegations into a "Rest of the World" category ensures statistical robustness.
Ultimately, any statistical uncertainty regarding whether a minor branch like Country 4 should have seen 4 or 6 departures is negligible for the broader global strategy.
The vast majority of the dataset is heavily concentrated in larger countries where the statistical validity is entirely robust.
Consequently, HR should acknowledge this limitation but can confidently rely on the macro-trends revealed by the table.

**Results Analysis: P-Value and Observed vs. Expected** The Pearson's Chi-squared test yields a $p$-value of 0.1269.
Since this is greater than the standard 0.05 threshold, the results are **not statistically significant** globally.
This implies that, overall, the differences in turnover rates across countries could simply be due to random chance rather than a systemic geographical issue.

However, looking at the absolute *Observed vs. Expected* differences, there are significant operational insights for specific branches:

-   **The Positive Outlier (Country 9):** Country 9 is not only the headquarters (accounting for over 1,000 employees), but its retention capacity is exceptionally strong. It retained almost 20 more employees than mathematically expected (`Dif_Stay` = 19.8). The company is saving talent and significant hiring costs in its most important market. HR should investigate what wellness, leadership, or compensation policies are being applied here to export them.
-   **The Critical Drain (Country 3):** Despite the global $p$-value, Country 3 requires attention. Its `Dif_Stay` is -6.1 (we expected 61 people to stay, but only 55 did). In the turnover column, we expected about 9 resignations, but they actually had 15. For a mid-sized office with a headcount of around 70, nearly doubling the expected number of resignations represents a critical talent drain.

### Interpretation: Should HR be concerned about country differences in individual turnover based on this analysis alone?

**No, not based on this analysis alone.** While the descriptive data shows concerning absolute numbers in branches like Country 3, HR should not launch expensive, country-specific retention initiatives based solely on this test for two main reasons:

-   **Lack of Statistical Significance:** The global $p$-value (0.1269) suggests that geography, as a standalone variable, is not a definitive driver of turnover across the whole company.
-   **Correlation Does Not Imply Causation:** This is a univariate analysis that does not tell the whole story. Country might just be a "proxy" for other underlying issues. For example, Country 3 might have higher turnover simply because it houses the Sales department (which naturally has higher turnover than Finance), or because its workforce has a lower average tenure or salary.

**Next Steps: Preview of promt 2** Before drawing definitive conclusions, HR needs to evolve from descriptive to predictive analytics.
We must build a multivariate model that controls for factors like department, salary, tenure, and performance score, to isolate whether the country is *truly* the root cause of the turnover problem.

# Prompt 2: Cross-Country Differences in Team Engagement and Turnover

### Tasks

-   Import 'Turnover_Team_Data.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Generate descriptive statistics (N, Mean, SD, Min, Max) for Team engagement and Team turnover rate by country.\
-   Conduct a one-way ANOVA for each variable: Test the assumption of homogeneity of variances using Levene’s test; Use Welch’s ANOVA if variances are unequal; Perform appropriate post-hoc tests to identify which countries differ.\

### Deliverables

-   Descriptive statistics tables.\
-   ANOVA output (including Welch correction if required).\
-   Post-hoc comparison table
-   Interpretation: Which country or countries should HR prioritize for further diagnosis, and why?

## Solution

### Descriptive statistics tables

```{r prompt2.1}
df2 <- read_excel("Databases/Turnover_Team_Data.xlsx")
str(df2)
head(df2)
df2$Country <- as.factor(df2$Country)
levels(df2$Country)
```


```{r prompt2.1}
library(dplyr)

df2 %>%
group_by(Country) %>%
summarise(
N = n(),
Eng_mean = mean(Engagement),
Eng_sd = sd(Engagement),
Eng_min = min(Engagement),
Eng_max = max(Engagement),
Turn_mean = mean(TeamSeparation),
Turn_sd = sd(TeamSeparation),
Turn_min = min(TeamSeparation),
Turn_max = max(TeamSeparation)
)
```
The dataset contains information on 212 teams across multiple countries. Each observation represents a team and includes variables related to team size, employee engagement, leadership quality, organizational culture, and turnover rates. The key variables for this analysis are team engagement and team turnover (TeamSeparation), while country is used as the grouping variable to test for cross-country differences.

### ANOVA output (including Welch correction if required)

```{r prompt2.2}
library(car)

#To see if we use normal ANOVA or Welch ANOVA
leveneTest(Engagement ~ Country, data = df2)
leveneTest(TeamSeparation ~ Country, data = df2)
```
Levene’s test was conducted to examine the assumption of homogeneity of variance across countries. The results were significant for both engagement (F = 10.805, p < 0.001) and team turnover (F = 5.368, p = 0.001), indicating that the assumption of equal variances was violated.
Consequently, Welch’s ANOVA was used instead of the standard ANOVA, as it does not assume equal variances across groups.

```{r prompt2.2}
oneway.test(Engagement ~ Country, data = df2)
oneway.test(TeamSeparation ~ Country, data = df2)
```
Welch’s ANOVA was conducted to examine whether engagement levels differ across countries. The results show a significant effect of country on engagement, F(3, 93.48) = 29.256, p < 0.001, indicating that engagement levels vary significantly between countries.
Welch’s ANOVA was also conducted to test whether team turnover differs across countries. The results indicate a significant effect of country on team turnover, F(3, 98.76) = 3.446, p = 0.020.

### Post-hoc comparison table

```{r prompt2.3}
install.packages("rstatix")
```


```{r prompt2.3}
library(rstatix)

games_howell_test(df2, Engagement ~ Country)
games_howell_test(df2, TeamSeparation ~ Country)
```
Post-hoc Games-Howell comparisons were conducted to identify which countries differed significantly. 
For engagement, Spain showed significantly lower engagement scores compared with the UK, US, and Canada. No significant differences were observed among the UK, US, and Canada.
For team turnover, Spain showed significantly higher turnover rates compared with the UK and the US. However, the difference between Spain and Canada was not statistically significant.

### Interpretation: Which country or countries should HR prioritize for further diagnosis, and why?
The results indicate that Spain should be prioritised for further diagnosis. Teams in Spain exhibit significantly lower engagement levels compared with the UK, US, and Canada, and also show higher team turnover rates. This combination suggests that underlying organisational or cultural factors may be affecting employee engagement and contributing to higher turnover in Spain. 
Consequently, HR should investigate potential drivers such as leadership practices, workload, work–life balance, or organisational culture in this country.


# Prompt 3: Individual-Level Predictors of Voluntary Turnover

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate a logistic regression model predicting leaver status using: Gender; Age; Length of service; Performance appraisal rating; Country (as a categorical variable).\
-   Evaluate overall model fit using: Likelihood ratio test; Nagelkerke R².\
-   Interpret coefficients using odds ratios.

### Deliverables

-   Logistic regression output.\
-   Table of log-odds and odds ratios.\
-   Interpretation: Which variables significantly predict turnover? How should HR interpret the effect of gender and performance ratings?

## Solution

### Logistic regression output

```{r prompt3.1}


```

### Table of log-odds and odds ratios

```{r prompt3.2}


```

### Interpretation: Which variables significantly predict turnover? How should HR interpret the effect of gender and performance ratings?

# Prompt 4: Gender Differences in Employee Turnover Over Tenure

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Conduct a Kaplan–Meier survival analysis using: Length of service as time variable; Leaver status as event indicator; Gender as grouping variable.\
-   Compare survival curves using a statistical test.\
-   Visualize survival probabilities over time.

### Deliverables

-   Survival table and summary statistics.\
-   Survival plot\
-   Interpretation: At what stages of tenure are gender differences in turnover most pronounced, and what HR processes might this implicate?

## Solution

### Survival table and summary statistics

```{r prompt4.1}


```

### Survival plot

```{r prompt4.2}


```

### Interpretation: At what stages of tenure are gender differences in turnover most pronounced, and what HR processes might this implicate?

# Prompt 5: Team-Level Drivers of Employee Turnover

### Tasks

-   Import 'Turnover_Team_Data.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate an OLS multiple regression model predicting team turnover using: Engagement; Team leader rating; CSR; Drive for performance; Performance, development and reward; Work–life balance; Country dummy variables (with one reference category).\
-   Examine: Model fit (R², F-test).\
-   Unstandardized and standardized coefficients.

### Deliverables

-   Regression output and Table of standardized coefficients.\
-   Interpretation: Which team-level factors are most strongly associated with turnover, and how should HR respond?

## Solution

### Regression output and Table of standardized coefficients

```{r prompt5.1}


```

### Interpretation: Which team-level factors are most strongly associated with turnover, and how should HR respond?

# Prompt 6: Team-Level Drivers of Employee Turnover

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate the proportion of individual turnover explained by gender alone using logistic regression.\
-   Use the Nagelkerke R² to approximate the share of turnover costs attributable to gender.\
-   Assume a total annual voluntary turnover cost of £4 million. Calculate potential cost savings if gender-related turnover were eliminated or reduced.

### Deliverables

-   Cost calculation.\
-   Executive-style summary answering: Is there a strong enough business case to justify a targeted retention intervention?
-   What cautions should HR keep in mind when interpreting these figures?

## Solution

### Cost calculation

```{r prompt6.1}


```

### Executive-style summary answering: Is there a strong enough business case to justify a targeted retention intervention?

### What cautions should HR keep in mind when interpreting these figures?
