---
# this is the Rmarkdown header. Use the YAML syntax. ---YAML block ---.
title: "Advanced Turnover Modelling and Business Case Development" 
subtitle: "Groupwork Use Case 3 - People Analytics"
author: "Ferrán Extremera, Irene Morales, Amalia Varo y Martina Iscar"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide

# the following lines tell R to store the output in the subdirectory reports
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
---

# Intro
Employee turnover is a critical concern for organizations due to its substantial financial and human costs. HR functions frequently report turnover statistics, but descriptive reporting alone may lead to misdiagnosis of problems and poorly targeted interventions.
In this use case, you will apply inferential and predictive analytics to understand voluntary employee turnover at both individual and team levels in a multinational organization

# Setup

Within the project directory, create the following folders if they do not already exist:

- **Databases**: to store all data sources provided in Moodle (Excel files).\
- **data**: to store any ad-hoc data created during the analysis. The native R data format is .rds.\
- **report**: to store the HTML output file generated when knitting the R Markdown file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(readxl)
```

# Prompt 1: Estimating differences in turnover by country

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Produce a cross-tabulation of Country × Leaver Status.\
-   Conduct a chi-square test of independence to assess whether turnover differs significantly by country. Interpret the results, paying attention to: expected vs observed frequencies; statistical significance;	limitations related to small subgroup sizes.\

### Deliverables

-   Cross-tabulation table with expected frequencies.\
-   Chi-square test output.\
-   Interpretation: Should HR be concerned about country differences in individual turnover based on this analysis alone?

## Solution
```{r}
df1 <- read_excel("Databases/Individual_Turnover.xlsx")
head(df1)
```
```{r}
str(df1)
```

```{r}
summary(df1)
```
Hay que tener cuidado, pues la variable Gender y LeaverStatus tienen 2 y 1 valor nulo. Por lo tanto, debemos quitarlos sobretodo por precaución y además al tratarse de como mucho tres filas del dataset frente a las 1653 filas totales, es tan solo una mínima eliminación que tiene como porpósito tener un dataset más puro y anticiparnos a posibles problemas que den estos valores nulos en los modelos usados a lo largo de la práctica.

```{r}
data <- df1 %>%
  drop_na(Gender, LeaverStatus)
```

```{r}
summary(data)
```
Comprobamos que se han eliminado las filas nulas y tenemos un total de 1650 datos para nuestro análisis.
```{r}
str(data)
```


### Cross-tabulation table with expected frequencies

```{r prompt1.1}
tabla_cruzada <- table(
  "País" = data$Country, 
  "Estado_Rotación (0=Retenido, 1=Salida)" = data$LeaverStatus
)

# Añadir una fila y columna de "Totales"
tabla_con_totales <- addmargins(tabla_cruzada)

# Imprimir en consola para que la veas tú mientras programas
cat("\n--- FRECUENCIAS OBSERVADAS ---\n")
print(tabla_con_totales)

# Crear la versión "bonita" para el reporte HTML final
knitr::kable(tabla_con_totales, 
             caption = "Frecuencia de Rotación de Empleados por País")
# Recordar cuando presentar poner el pais correspondiente al lado de número 
```

### Chi-square test output

```{r prompt1.2}
# 1. Asegurarnos de tener el test calculado
test_chi <- chisq.test(tabla_cruzada)
```
```{r}
cat("\n--- RESULTADO DEL TEST CHI-CUADRADO ---\n")
print(test_chi)
```
```{r}
print(test_chi$expected)
```

```{r prompt1.2}
# 2. Crear una tabla completa combinando las columnas del test
# La columna 1 es "Se quedan" (0) y la columna 2 es "Se van" (1)
tabla_completa <- data.frame(
  País = rownames(test_chi$observed),
  
  # Datos REALES (Lo que de verdad pasó)
  Stay_obv = test_chi$observed[, 1],
  Leave_obv    = test_chi$observed[, 2],
  
  # Datos ESPERADOS (Lo que la estadística predecía si todo fuera equitativo)
  Stay_exp = round(test_chi$expected[, 1], 1),
  Leave_exp    = round(test_chi$expected[, 2], 1)
)

tabla_completa$Dif_Stay <- tabla_completa$Stay_obv - tabla_completa$Stay_exp

# 3. Mostrar la tabla con un diseño elegante para tu reporte HTML
knitr::kable(tabla_completa, 
             row.names = FALSE, 
             caption = "Complete Comparison: Retention and Turnover (Observed vs Expected) by Country")

```

### Interpretation: Should HR be concerned about country differences in individual turnover based on this analysis alone?
Interpretation: Should HR be concerned about country differences in individual turnover based on this analysis alone?
(Nota: Como no tengo los resultados numéricos reales, te dejo la plantilla de interpretación con los insights que debes rellenar según lo que te arroje el código anterior).

1. Diferencias entre lo Esperado y lo Observado:

Insight: Al comparar la tabla de "Frecuencias Observadas" con la de "Frecuencias Esperadas", podemos identificar qué países están impulsando la rotación. Si un país tiene una frecuencia observada de salidas (Leavers) mucho mayor que su frecuencia esperada, es un foco rojo. Esto indica que ese país está perdiendo talento a un ritmo superior al promedio general de la empresa.

2. Significancia Estadística:

Insight: Si el p-value del test es menor a 0.05, podemos afirmar que existe una relación estadísticamente significativa entre el país y la decisión de irse de la empresa (la rotación no es aleatoria respecto a la geografía). Si el p-value es mayor a 0.05, las diferencias observadas podrían deberse simplemente al azar.

3. Limitaciones por Tamaños de Subgrupo (Small Subgroup Sizes):

Insight: El test de Chi-cuadrado asume que tenemos suficientes datos. Si en la tabla de frecuencias esperadas vemos valores menores a 5 (por ejemplo, en países donde la empresa tiene operaciones muy pequeñas o casi nadie ha renunciado), el cálculo matemático pierde precisión y podría arrojar falsas alarmas.

Conclusión para HR (¿Deberían preocuparse basándose solo en este análisis?):
No de forma aislada. Aunque el test muestre diferencias significativas, este es un análisis univariado que no cuenta toda la historia.

Correlación no implica causalidad: El país podría ser solo un "proxy" de otros problemas. Por ejemplo, tal vez un país tiene mayor rotación simplemente porque concentra al equipo de Ventas (que por naturaleza tiene más rotación que Finanzas), o porque su plantilla es mucho más joven.

Siguientes pasos recomendados para HR: Antes de lanzar costosas iniciativas de retención por país, HR necesita un modelo predictivo multivariable (como una regresión logística) que controle factores como el departamento, el salario, la antigüedad o el desempeño, para aislar si el país es realmente la causa del problema.




1. ¿Cuál es el problema exactamente? (La Regla del 5)
Para que el cálculo matemático del test de Chi-cuadrado sea fiable, tiene una regla estricta: Las frecuencias ESPERADAS (no las reales, sino las calculadas por la estadística) en cada "casilla" de tu tabla deben ser mayores a 5. Si miras tu tabla de "Frecuencias Esperadas" y ves números como 2.3, 4.1 o 0.8, tienes un problema de tamaño de muestra.

2. ¿Por qué ocurre esto en Recursos Humanos?
Imagina que la empresa tiene una oficina muy pequeña en el País 10, con solo 6 empleados en total. Si la tasa de rotación global de la empresa es del 10%, la estadística "esperaría" que en ese país se fueran 0.6 personas (el 10% de 6).

Como 0.6 es menor que 5, el test de Chi-cuadrado se "marea". La fórmula matemática no está diseñada para manejar grupos tan pequeños y pierde precisión.

3. ¿Cómo te avisa R de este problema?
Si intentas hacer el chisq.test() con grupos muy pequeños, R te va a lanzar un mensaje de advertencia (Warning) en la consola que dice exactamente esto:

Warning message: In chisq.test(tabla_cruzada) : Chi-squared approximation may be incorrect (La aproximación de Chi-cuadrado puede ser incorrecta).

4. ¿Qué impacto tiene para la empresa (El Insight)?
Si ignoras esta advertencia, podrías darle un falso positivo a Recursos Humanos. Podrías decirles: "¡Alerta roja! El País 10 tiene una rotación significativamente alta"... cuando en realidad, solo se fue una persona, pero como la oficina tiene solo 4 empleados, pareció un porcentaje altísimo (25%). Estarías haciendo que HR gaste tiempo y dinero en investigar un país que no tiene un problema real, es solo ruido estadístico.

# Prompt 2: Cross-Country Differences in Team Engagement and Turnover

### Tasks

-   Import 'Turnover_Team_Data.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Generate descriptive statistics (N, Mean, SD, Min, Max) for Team engagement and Team turnover rate by country.\
-   Conduct a one-way ANOVA for each variable: Test the assumption of homogeneity of variances using Levene’s test; Use Welch’s ANOVA if variances are unequal; Perform appropriate post-hoc tests to identify which countries differ.\


### Deliverables

-   Descriptive statistics tables.\
-   ANOVA output (including Welch correction if required).\
-   Post-hoc comparison table
-   Interpretation: Which country or countries should HR prioritize for further diagnosis, and why?

## Solution

### Descriptive statistics tables

```{r prompt2.1}


```

### ANOVA output (including Welch correction if required)

```{r prompt2.2}


```

### Post-hoc comparison table

```{r prompt2.3}


```

### Interpretation: Which country or countries should HR prioritize for further diagnosis, and why?

# Prompt 3: Individual-Level Predictors of Voluntary Turnover

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate a logistic regression model predicting leaver status using: Gender; Age; Length of service; Performance appraisal rating; Country (as a categorical variable).\
-   Evaluate overall model fit using: Likelihood ratio test; Nagelkerke R².\
-   Interpret coefficients using odds ratios.

### Deliverables

-   Logistic regression output.\
-   Table of log-odds and odds ratios.\
-   Interpretation: Which variables significantly predict turnover? How should HR interpret the effect of gender and performance ratings?

## Solution

### Logistic regression output

```{r prompt3.1}


```

### Table of log-odds and odds ratios

```{r prompt3.2}


```

### Interpretation: Which variables significantly predict turnover? How should HR interpret the effect of gender and performance ratings?

# Prompt 4: Gender Differences in Employee Turnover Over Tenure

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Conduct a Kaplan–Meier survival analysis using: Length of service as time variable; Leaver status as event indicator; Gender as grouping variable.\
-   Compare survival curves using a statistical test.\
-   Visualize survival probabilities over time.

### Deliverables

-   Survival table and summary statistics.\
-   Survival plot\
-   Interpretation: At what stages of tenure are gender differences in turnover most pronounced, and what HR processes might this implicate?

## Solution

### Survival table and summary statistics

```{r prompt4.1}


```

### Survival plot

```{r prompt4.2}


```

### Interpretation: At what stages of tenure are gender differences in turnover most pronounced, and what HR processes might this implicate?

# Prompt 5: Team-Level Drivers of Employee Turnover

### Tasks

-   Import 'Turnover_Team_Data.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate an OLS multiple regression model predicting team turnover using: Engagement; Team leader rating; CSR; Drive for performance; Performance, development and reward; Work–life balance; Country dummy variables (with one reference category).\
-   Examine: Model fit (R², F-test).\
-   Unstandardized and standardized coefficients.

### Deliverables

-   Regression output and Table of standardized coefficients.\
-   Interpretation: Which team-level factors are most strongly associated with turnover, and how should HR respond?

## Solution

### Regression output and Table of standardized coefficients

```{r prompt5.1}


```

### Interpretation: Which team-level factors are most strongly associated with turnover, and how should HR respond?

# Prompt 6: Team-Level Drivers of Employee Turnover

### Tasks

-   Import 'Individual_Turnover.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Estimate the proportion of individual turnover explained by gender alone using logistic regression.\
-   Use the Nagelkerke R² to approximate the share of turnover costs attributable to gender.\
-   Assume a total annual voluntary turnover cost of £4 million. Calculate potential cost savings if gender-related turnover were eliminated or reduced.

### Deliverables

-   Cost calculation.\
-   Executive-style summary answering: Is there a strong enough business case to justify a targeted retention intervention?
-   What cautions should HR keep in mind when interpreting these figures?

## Solution

### Cost calculation

```{r prompt6.1}


```

### Executive-style summary answering: Is there a strong enough business case to justify a targeted retention intervention?

### What cautions should HR keep in mind when interpreting these figures?
