---
title: "Recruitment and Selection Analytics" 
subtitle: "Groupwork Use Case 5 - People Analytics"
author: "Substitute Group_X by your Group number in the name of the .Rmd file"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
editor_options: 
  markdown: 
    wrap: 72
---

# Intro

In this final workshop, you will apply analytics techniques to
recruitment and selection data to identify potential biases, evaluate
the fairness of selection decisions, and assess whether assessment
centre data can predict performance and turnover. You will use
chi-square tests, logistic and linear regression models to uncover
hidden patterns and build evidence-based recommendations.

# Setup

Within the project directory, create the following folders if they do
not already exist:

-   **Databases**: to store all data sources provided in Moodle (Excel
    files).\
-   **data**: to store any ad-hoc data created during the analysis. The
    native R data format is .rds.\
-   **report**: to store the HTML output file generated when knitting
    the R Markdown file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
```

# Prompt 1: Exploring Diversity in the Applicant Pool

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Recode gender and UG variables.\
-   Generate frequency tables for Gender, UG, Shortlisted, Interviewed,
    Offered, Accepted, Joined.\
-   Interpret patterns in diversity at each recruitment stage.\

### Deliverables

-   Table of frequencies.\
-   Initial diagnosis: Where in the pipeline are UG or female applicants
    being lost?

## Solution

### Table of frequencies

```{r prompt1.1}
# Load libraries
library(readxl)
library(dplyr)
library(janitor)

# Import dataset
df1 <- read_excel("Databases/recruitment_applicants_2023.xlsx")

# Clean names
df1 <- clean_names(df1)

# Check names
names(df1)
names(df1)
glimpse(df1)
# Recode variables
# Recode variables
df1 <- df1 %>%
  mutate(
    gender_rec = case_when(
      gender %in% c(1, "1", "F", "Female") ~ "Female",
      gender %in% c(2, "2", "M", "Male") ~ "Male",
      TRUE ~ "Other/Unknown"
    ),
    ug_rec = case_when(
      u_gyn %in% c(1, "1", "Y", "Yes") ~ "UG",
      u_gyn %in% c(2, "2", "N", "No") ~ "Non-UG",
      TRUE ~ "Unknown"
    )
  )

# Frequency tables
freq_gender <- tabyl(df1$gender_rec)
freq_ug <- tabyl(df1$ug_rec)
freq_shortlisted <- tabyl(df1$shortlisted_ny)
freq_interviewed <- tabyl(df1$interviewed)
freq_offered <- tabyl(df1$offer_ny)
freq_accepted <- tabyl(df1$accept_ny)
freq_joined <- tabyl(df1$join_yn)

freq_gender
freq_ug
freq_shortlisted
freq_interviewed
freq_offered
freq_accepted
freq_joined

# Print results

```

The results provide a clear overview of how the recruitment process
operates and where the main issues may lie.

First, the applicant pool is already imbalanced in terms of gender, with
only around 28% female candidates compared to over 70% male candidates.
This suggests that part of the diversity challenge originates before the
selection process itself, likely at the attraction or application stage.
In contrast, the proportion of candidates from underrepresented groups
is more balanced, which makes it possible to meaningfully assess whether
the selection process treats these groups fairly.

Looking at the progression through the recruitment stages, the data
shows a strong filtering effect in the early phases. Only about 31% of
applicants are shortlisted, and less than 20% reach the interview stage.
This indicates that the majority of candidates are eliminated very early
in the process. However, once candidates reach the interview stage, the
process becomes more selective but also more efficient, with
approximately half of interviewed candidates receiving an offer.
Finally, around 6% of the original applicant pool ends up accepting and
joining the organization.

This pattern suggests that the main bottleneck occurs during the
shortlisting and interview selection phases. These stages are typically
more subjective and involve greater human judgment, which increases the
risk of inconsistency and potential bias. According to recruitment
analytics theory, such stages are particularly vulnerable to effects
like first impressions or unconscious stereotyping, which may influence
decisions even when formal criteria are in place .

Another important aspect is the structure of the data itself. The large
number of missing values in later stages (offers, acceptance, and
joining) is expected, as only candidates who progress through earlier
stages are observed in subsequent ones. However, this means that any
deeper analysis must account for this conditional structure, otherwise
the conclusions could be misleading.

Overall, the results indicate that the recruitment process is highly
selective at the beginning and more consistent in later stages. While
this may reflect an efficient filtering mechanism, it also raises
concerns about whether early decisions are being made fairly and
systematically. Further statistical analysis, such as chi-square tests
or regression models, is necessary to determine whether these patterns
reflect true merit-based selection or potential bias against certain
groups.
### Initial diagnosis: Where in the pipeline are UG or female
applicants being lost? The data indicates that the largest drop in
candidates occurs during the transition from application to
shortlisting, and then again before the interview stage. Since these are
the most restrictive phases of the process, they are also the most
likely points where female or UG applicants could be disproportionately
filtered out.

Although the current results do not yet provide a breakdown by gender or
UG at each stage, the structure of the pipeline strongly suggests that
any loss of diversity is happening early rather than later. Once
candidates reach the interview stage, progression rates improve
significantly, making it less likely that major disparities occur at
that point.

Therefore, the initial diagnosis is that shortlisting is the critical
stage where potential bias or unequal outcomes may arise. Further
analysis, such as comparing selection rates by gender and UG, would be
necessary to confirm whether these groups are being systematically
disadvantaged

# Prompt 2: Testing for Bias in Shortlisting and Offers

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Use cross-tabulation and chi-square tests to compare: Gender ×
    Shortlisted; UG × Shortlisted; Gender × Offer; UG × Offer.\
-   Identify significant imbalances.\

### Deliverables

-   Chi-square test results with interpretation.\
-   Summary: Are the recruitment decisions fair and unbiased?\

## Solution

### Chi-square test results with interpretation

```{r prompt2.1}
library(readxl)
df2 <- read_excel("Databases/recruitment_applicants_2023.xlsx")
```

```{r}
names(df2)
```

```{r}
str(df2)
```

```{r prompt2.2}
# Quick checks
summary(df2)
```

```{r prompt2.3}
# 2. Recodificar las variables relevantes para que las tablas sean legibles
df2 <- df2 %>%
  mutate(
    Gender = factor(Gender, levels = c(1, 2), labels = c("Male", "Female")),
    UGyn = factor(UGyn, levels = c(1, 2), labels = c("Yes", "No")),
    ShortlistedNY = factor(ShortlistedNY, levels = c(0, 1), labels = c("Rejected", "Shortlisted")),
    OfferNY = factor(OfferNY, levels = c(0, 1), labels = c("No Offer", "Made Offer"))
  )
```

```{r prompt2.4}
# 3. Generar Tablas Cruzadas y Pruebas Chi-cuadrado
# Gender × Shortlisted
cat("\n--- Gender x Shortlisted ---\n")
tab_gen_short <- table(df2$Gender, df2$ShortlistedNY)
print(tab_gen_short)
test_chi_gen <- chisq.test(tab_gen_short)
print(test_chi_gen)
```

```{r}
print(test_chi_gen$expected)
```

The analysis of the shortlisting stage aimed to evaluate the fairness and potential existence of gender bias in the initial recruitment decisions. The Chi test determines whether a statistically significant relationship exists between the candidates' gender and their probability of passing this first evaluation filter.

The results yielded a p-value of 0.00019 \< 0.05, it conclusively confirms that shortlisting decisions are not independent of the applicant's gender. When contrasting this finding with the actual proportions of the sample, the disparity becomes evident: while almost half of the male candidates (48.7%) managed to advance to the interview phase, barely a quarter of the female candidates (24.8%) made the cut.

Therefore, the current shortlisting process is not impartial and presents a statistically significant bias favoring men. Despite the organization attracting a volume of female talent that almost triples the male talent in the application phase (202 women compared to 78 men), their rejection rate is disproportionately high. This data points to a clear "bottleneck" for diversity at this early stage, suggesting an urgent need to audit and recalibrate the criteria or algorithms used during the initial profile review.

```{r prompt2.5}
# UG × Shortlisted
cat("\n--- UG (Underrepresented Group) x Shortlisted ---\n")
tab_ug_short <- table(df2$UGyn, df2$ShortlistedNY)
print(tab_ug_short)
test_chi_ug <-chisq.test(tab_ug_short)
print(test_chi_ug)
```

```{r}
print(test_chi_ug$expected)
```

Additionally in the analysis of the shortlisting stage we evaluated the fairness of decisions regarding candidates belonging to underrepresented groups (UG).

The statistical results give a p-value of **1.472e-06**, the value is drastically lower than the standard significance level of 0.05, it confirms that a highly significant statistical relationship exists between the two variables. This means that the decision to shortlist a candidate is not independent of their underrepresented group status; on the contrary, there is a clear, non-random pattern in the results.

When contrasting the observed proportions, the disadvantage is evident. Only **15.7%** of the candidates belonging to underrepresented groups (19 out of 121) managed to advance to the next stage. In contrast, candidates who do not belong to these groups had a much higher success rate, reaching a **43.4%** shortlisting rate (69 out of 159).

This disparity is conclusively confirmed when reviewing the table of expected values. In a scenario where the selection process is completely neutral and equitable, statistics project that approximately 38 candidates from underrepresented groups should have been shortlisted. However, in reality, only 19 managed to pass the filter—exactly half of what would be expected under fair conditions.

The shortlisting process also presents a significant bias against underrepresented groups. The disproportionate rejection rate toward this demographic and the large deviation from the expected values indicate that the criteria or methods used to filter candidates are not impartial.

```{r prompt2.6}
# Gender × Offer
cat("\n--- Gender x Offer ---\n")
tab_gen_offer <- table(df2$Gender, df2$OfferNY)
print(tab_gen_offer)
test_chi_gen_off <- chisq.test(tab_gen_offer)
print(test_chi_gen_off)
```

```{r}
print(test_chi_gen_off$expected)
```

Moving on to teh next stage, we evaluated the fairness with the same variables, starting with gender.The aim is to determine whether the probability of a candidate receiving a formal proposal is independent of their gender.

Because the p-value is 0.04279, it confirms that a statistically significant relationship exists between the two variables. This indicates that hiring decisions in this phase are not random and present an imbalance dependent on the candidate's gender.

When analyzing the actual proportions of the sample, the bias becomes evident. Out of the 27 men who reached this stage, a large majority of 66.7% (18 candidates) received a job offer. In stark contrast, out of the 28 women evaluated, only 35.7% (10 candidates) managed to secure a proposal. This disparity is reaffirmed by examining the table of expected values: in a scenario of total equity, statistics projected that both groups would receive approximately 14 offers each (13.75 for men and 14.25 for women). However, in practice, men far exceeded this expectation, while women fell significantly below it.

In conclusion, the quantitative evidence demonstrates that the final offer stage is not impartial and presents a statistically significant gender bias. Despite having reached the same evaluative stage, male candidates have a considerably higher probability of being hired than their female counterparts, suggesting that the decision criteria in this final phase require a review to ensure fairness.

```{r prompt2.7}
# UG × Offer
cat("\n--- UG (Underrepresented Group) x Offer ---\n")
tab_ug_offer <- table(df2$UGyn, df2$OfferNY)
print(tab_ug_offer)
test_chi_ug_off <- chisq.test(tab_ug_offer)
print(test_chi_ug_off)
```

```{r}
print(test_chi_ug_off$expected)
```

Finally doing the same statistical analysis of the final stage of the selection process with regard to belonging to an underrepresented group, seeking to determine whether underrepresented group status influences the probability of a finalist receiving a formal hiring proposal.

The results of the statistical test yielded a p-value of 0.5756. Since this indicator is considerably higher than the standard significance level of 0.05, there is no statistical evidence to assert that a relationship exists between the two variables. This means that the decisions made in this phase are independent of whether or not the candidate belongs to an underrepresented group; the observed differences are purely due to normal random variance and not to a systemic bias.

When analyzing the proportions and the expected values, this fairness is confirmed. Out of the 13 candidates from underrepresented groups who reached this stage, 8 received an offer (61.5%). On the other hand, out of the 42 candidates from non-underrepresented groups, 20 received a proposal (47.6%). The actual figures align very closely with the statistical projections of a perfectly neutral scenario: the model expected that approximately 6.6 candidates from underrepresented groups would receive an offer, and in practice, there were 8.

In conclusion, the quantitative evidence indicates that the final offer stage is fair and impartial regarding underrepresented groups. The data demonstrates that once these candidates manage to reach the final phase of the process, they do not face a discriminatory barrier to being hired, as their profiles are evaluated equitably compared to the rest of the applicants. But with the previous analysis just to reach this stage there is a bias regarding if you belong to the underrepresented group or not.

### Summary: Are the recruitment decisions fair and unbiased?

Based on the quantitative analysis, the recruitment decisions are **not completely fair or unbiased**. While there is evidence of equitable treatment in one specific instance, significant systemic biases exist—particularly against women throughout the entire pipeline and against underrepresented groups during the initial screening.

Here is a breakdown of the fairness across the different stages:

#### **1. The Shortlisting Stage: A Major Diversity Bottleneck**

The initial screening phase (*shortlisting*) lacks objectivity and presents severe, statistically significant biases against both women and underrepresented groups.

-   **Gender Bias:** The process disproportionately favors male candidates ($p = 0.00019$). Although the organization attracted nearly three times as many female applicants (202 women vs. 78 men), their rejection rate was massive. Only **24.8%** of female candidates advanced, compared to nearly half (**48.7%**) of the male candidates.

-   **Underrepresented Groups (UG) Bias:** The shortlisting phase also heavily penalizes UG talent ($p = 0.00000147$). Only **15.7%** of UG candidates successfully passed the initial filter, whereas non-UG candidates enjoyed a **43.4%** success rate. In a perfectly equitable scenario, statistical models expected around 38 UG candidates to advance; in reality, only exactly half of that number (19) succeeded.

#### **2. The Offer Stage: Mixed Outcomes**

At the final hiring stage, the outcomes diverge depending on the demographic group being evaluated.

-   **Persistent Gender Bias:** The bias against women continues into the final decision-making process ($p = 0.04279$). Among the finalists, a sweeping **66.7%** of men received a formal job offer, compared to a mere **35.7%** of women. While statistical projections expected an even split (approximately 14 offers for each gender), male candidates significantly exceeded expectations at the direct expense of female candidates.

-   **Fairness for Underrepresented Groups:** Conversely, the final offer stage is completely fair and impartial regarding UG status ($p = 0.5756$). The data shows no systemic barrier for UG candidates who make it to this phase. In fact, **61.5%** of UG finalists received an offer (compared to 47.6% for non-UG candidates), which perfectly aligns with the expected values of a neutral and unbiased selection scenario.

Overall, the recruitment process is **not fair**. The data irrefutably points to a highly discriminatory initial shortlisting process that arbitrarily filters out diverse talent, followed by a final decision phase that continues to disadvantage women. The organization urgently needs to audit and recalibrate the criteria, algorithms, and human judgments used in its recruitment pipeline to guarantee true equity and inclusion.

Next we provide with graphsx to visually interpret the bias:

```{r visual_gender_pipeline, fig.width=8, fig.height=5}
# Asegúrate de tener cargado tidyverse
library(tidyverse)

# 1. Preparar los datos del pipeline directamente desde df2
pipeline_gender <- df2 %>%
  group_by(Gender) %>%
  summarise(
    `1. Applicants` = n(),
    `2. Shortlisted` = sum(ShortlistedNY == "Shortlisted", na.rm = TRUE),
    `3. Offered` = sum(OfferNY == "Made Offer", na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -Gender, names_to = "Stage", values_to = "Count")

# 2. Crear el gráfico de caída (Pipeline)
ggplot(pipeline_gender, aes(x = Stage, y = Count, group = Gender, color = Gender)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 4) +
  geom_text(aes(label = Count), vjust = -1.5, size = 5, show.legend = FALSE) +
  scale_color_manual(values = c("Male" = "#2c3e50", "Female" = "#e74c3c")) + # Azul oscuro y Rojo
  theme_minimal() +
  labs(
    title = "Recruitment Pipeline by Gender: The Diversity Bottleneck",
    subtitle = "While attracting almost 3x more women, the female talent pool drops drastically at the shortlisting stage.",
    x = "Recruitment Stage",
    y = "Number of Candidates",
    color = "Gender"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(size = 11, face = "bold"),
    legend.position = "top"
  )
```

```{r visual_ug_bias, fig.width=8, fig.height=5}
# 1. Extraer los datos exactos de la prueba Chi-cuadrado que ya hiciste
test_chi_ug <- chisq.test(table(df2$UGyn, df2$ShortlistedNY))

# 2. Construir un dataframe resumen para el gráfico
df_obs_exp_ug <- data.frame(
  Group = c("UG (Underrepresented)", "UG (Underrepresented)", "Non-UG", "Non-UG"),
  Metric = factor(c("Actual (Observed)", "Fair Expectation (Expected)", 
                    "Actual (Observed)", "Fair Expectation (Expected)"),
                  levels = c("Fair Expectation (Expected)", "Actual (Observed)")),
  Count = c(
    test_chi_ug$observed["Yes", "Shortlisted"],
    test_chi_ug$expected["Yes", "Shortlisted"],
    test_chi_ug$observed["No", "Shortlisted"],
    test_chi_ug$expected["No", "Shortlisted"]
  )
)

# 3. Crear el gráfico de barras agrupadas
ggplot(df_obs_exp_ug, aes(x = Group, y = Count, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = round(Count, 0)), 
            position = position_dodge(width = 0.8), vjust = -0.5, size = 5, fontface = "bold") +
  scale_fill_manual(values = c("Fair Expectation (Expected)" = "#bdc3c7", # Gris neutro
                               "Actual (Observed)" = "#e74c3c")) +         # Rojo alerta
  theme_minimal() +
  labs(
    title = "Shortlisting Bias: Actual Hires vs. Fair Expectations",
    subtitle = "Underrepresented groups (UG) fall 50% short of what a completely unbiased process would predict.",
    x = "Candidate Group",
    y = "Number of Shortlisted Candidates",
    fill = ""
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(size = 12, face = "bold"),
    legend.position = "top"
  )
```

# Prompt 3: Predicting Shortlisting Using Demographics

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Run logistic regression with Gender and UG as predictors and
    ShortlistedNY as target.\
-   Interpret odds ratios and model fit (Nagelkerke R²).\
-   Comment on potential compound bias (e.g., female & UG status).

### Deliverables

-   Model output.\
-   Odds ratios and statistical significance.\
-   Recommendations to improve fairness in shortlisting.\

## Solution

### Model output

```{r prompt3.1}
# Cargar la librería necesaria para calcular el Nagelkerke R²
if(!require(DescTools)) install.packages("DescTools")
library(DescTools)

# Para la regresión logística, es más seguro usar una variable objetivo numérica (0 y 1).
# Convertimos 'ShortlistedNY' (1 = Shortlisted, 0 = Rejected).
df2$Shortlisted_num <- ifelse(df2$ShortlistedNY == "Shortlisted", 1, 0)

# Ajustar el modelo de regresión logística con Gender y UGyn como predictores
model_shortlist <- glm(Shortlisted_num ~ Gender + UGyn, 
                       data = df2, 
                       family = binomial)

# Imprimir el resumen del modelo para ver los coeficientes y los p-valores (significancia estadística)
summary(model_shortlist)

# Calcular y mostrar el ajuste del modelo (Nagelkerke R²)
cat("\n--- Model Fit (Nagelkerke R²) ---\n")
PseudoR2(model_shortlist, which = "Nagelkerke")

```

### Odds ratios and statistical significance

```{r prompt3.2}
# Calcular los Odds Ratios exponencializando los coeficientes del modelo
cat("\n--- Odds Ratios ---\n")
odds_ratios <- exp(coef(model_shortlist))
print(odds_ratios)

# Calcular los intervalos de confianza al 95% para los Odds Ratios
cat("\n--- 95% Confidence Intervals for Odds Ratios ---\n")
conf_intervals <- exp(confint(model_shortlist))
print(conf_intervals)
```

### Recommendations to improve fairness in shortlisting

Both gender and underrepresented group status are highly significant predictors in the shortlisting model, as their p-values are well below 0.05. This provides strong statistical evidence that the initial screening process is not demographically neutral.

The model yields a Nagelkerke $R^{2}$ of 19.55%. While a 19.55% explanatory power might seem relatively low in some statistical contexts, in People Analytics and specifically when evaluating recruitment fairness, this is a very substantial and concerning figure. It indicates that nearly one-fifth of the variance in shortlisting decisions is driven solely by a candidate's gender and UG status, rather than their actual skills or merit.

-   **Gender Disadvantage:** the negative coefficient for female candidates (-1.19) translates to an odds ratio of approximately 0.30. This means that, holding other factors constant, women have 70% lower odds of being shortlisted compared to men. The data clearly shows that being a woman in this specific selection process is a major disadvantage.

-   **UG Status Advantage for the Majority:** Not belonging to an underrepresented group yields a large positive coefficient (+1.51), which translates to an odds ratio of approximately 4.55. This indicates that majority candidates (Non-UG) have 4.5 times higher odds of passing the screening phase compared to candidates from underrepresented groups.

**Conclusion & Recommendations**

The model reveals a severe systemic bias in the shortlisting phase. Furthermore, the combination of these factors points to a severe compound bias: a candidate who is both female and belongs to an underrepresented group faces a multiplicative penalty, making her chances of being shortlisted statistically minimal compared to a non-UG male candidate. To mitigate this, the organization should strongly consider implementing blind CV screening and structured evaluation rubrics for the initial stages of recruitment.

# Prompt 4: Validating Selection Tools Using Performance Ratings

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are
    included in the 'variable key' worksheet within the Excel file).\
-   Predict Year 1 performance rating using demographic, personality,
    competency, aptitude, and onboarding variables.\
-   Use linear regression with standardized coefficients.\
-   Interpret model fit and variable importance.

### Deliverables

-   Regression summary with top predictors.\
-   Evaluation of selection validity: Which tools truly predict
    performance?\

## Solution

### Regression summary with top predictors

```{r prompt4.1}
library(readxl)
library(dplyr)
library(lm.beta)
library(broom)
library(car)

# Import data
selection <- read_excel("Databases/selection.xlsx", sheet = "data")

# Keep only observations with Year 1 performance available
selection_model <- selection %>%
  filter(!is.na(Year1performanceRating))

# Linear regression
# Omit HRDummyV so HR becomes the reference category
model_p4 <- lm(
  Year1performanceRating ~
    Gender + EducationHighest + UGYN + WorkExperience +
    ACPersonalityO + ACPersonalityC + ACPersonalityE + ACPersonalityA + ACPersonalityN +
    ACRatingINTCOMPA + ACRatingINTCOMPB + ACRatingINTCOMPC +
    ACRatingINTCOMPD + ACRatingINTCOMPE +
    ACRatingAPTnumerical + ACRatingAPTverbal +
    InductionDay + InductionWeek + OnBoardingBuddy +
    FinanceDummyV + MarketingDummyV + SalesDummyV +
    RiskDummyV + LegalDummyV + OperationsDummyV,
  data = selection_model
)

# Standardized coefficients
model_p4_beta <- lm.beta(model_p4)

# Full summary
summary(model_p4_beta)

# Tidy table with standardized coefficients
p4_results <- broom::tidy(model_p4_beta) %>%
  mutate(std_beta = c(NA, lm.beta(model_p4)$standardized.coefficients[-1])) %>%
  arrange(desc(abs(std_beta)))

p4_results

# Optional: multicollinearity check
car::vif(model_p4)

```

**Leyend:** *ACRatingINTCOMPA = Assessment Centre Competency, Technical
ACRatingINTCOMPB = Assessment Centre Competency, Team Player
ACRatingINTCOMPC = Assessment Centre Competency Critical Thought
ACRatingINTCOMPD = Assessment Centre Competency Business Aware,
ACRatingINTCOMPE = Assessment Centre Competency Drive Innovation*

A multiple linear regression was conducted to predict **Year 1
performance rating** from demographic variables, personality measures,
assessment-centre competency ratings, aptitude test scores, onboarding
variables, and job-function controls. The overall model was
statistically significant, **F(25, 292) = 23.08, p \< .001**, indicating
that the predictors jointly explained a meaningful proportion of
variance in performance outcomes. The model showed strong explanatory
power, with **R² = 0.664** and **adjusted R² = 0.635**, meaning that
around 64% of the variance in Year 1 performance ratings was accounted
for by the included variables.

Examining the standardized coefficients, the strongest significant
predictors in the full model were **Gender** (β = -0.362, p \< .001),
**Legal job function** (β = 0.357, p \< .001), **Marketing job
function** (β = 0.224, p = .005), **Risk job function** (β = 0.219, p \<
.001), **Assessment Centre Personality Extraversion** (β = 0.179, p =
.006), **Assessment Centre Competency Technical** (β = 0.155, p = .009),
**Assessment Centre Competency Business Aware** (β = 0.150, p = .002),
**Assessment Centre Personality Conscientious** (β = 0.144, p = .003),
**Assessment Centre Competency Drive Innovation** (β = 0.123, p = .038),
and **Attend Induction Day** (β = 0.102, p = .012). In contrast, **UG
status**, **work experience**, **Openness**, **Agreeableness**,
**Neuroticism**, **Team Player**, **Critical Thought**, **Numerical
Aptitude Test Score**, **Verbal Reasoning Aptitude Test Score**,
**Induction Week**, and **OnBoardingBuddy** were not statistically
significant predictors in the full model.

### Evaluation of selection validity: Which tools truly predict performance?

To answer the validity question properly, it is important to distinguish
between **contextual control variables** and actual **selection tools**.
Although **Gender** and several **job-function dummies** were among the
strongest predictors in the model, these variables do not represent
selection methods themselves. Instead, they indicate that performance
ratings vary across demographic coding and business areas, so they
should be interpreted as contextual influences rather than evidence of
selection validity.

Focusing specifically on the selection-related measures, the clearest
evidence of predictive validity comes from selected **personality** and
**competency** indicators. In particular, **Extraversion** and
**Conscientiousness** were both positive and significant predictors of
Year 1 performance, suggesting that these traits are meaningfully
associated with stronger early job outcomes. In addition, the competency
ratings for **Technical**, **Business Aware**, and **Drive Innovation**
all made significant positive contributions, indicating that these
assessment-centre dimensions successfully capture capabilities linked to
later performance. Among onboarding measures, only **Attend Induction
Day** showed a significant effect, while **Induction Week** and
**OnBoardingBuddy** did not.

Overall, the findings suggest that the organization’s selection system
is **partially valid rather than uniformly valid**. Some tools clearly
contribute to predicting future performance, especially selected
personality dimensions and specific competency ratings. However, other
measures appear to add little incremental value once the full model is
taken into account. Most notably, the **numerical** and **verbal
aptitude tests** were not significant predictors in this analysis, nor
were the competency dimensions **Team Player** and **Critical Thought**.
This implies that the organization should place greater emphasis on the
measures that demonstrate clear predictive value, while critically
reviewing whether weaker tools meaningfully improve hiring decisions. In
short, the results support the validity of parts of the assessment
process, but they do not justify treating the entire assessment battery
as equally effective.

# Prompt 5: Predicting Early Turnover

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are
    included in the 'variable key' worksheet within the Excel file).\
-   Run logistic regression to predict LeaverYr2.\
-   Use demographics, assessments, onboarding, and performance rating.\
-   Evaluate odds ratios and risk factors.

### Deliverables

-   Model output and interpretation.\
-   Recommendations for reducing graduate attrition.\

## Solution

### Model output and interpretation

```{r prompt5.1}
library(readxl)
library(dplyr)
library(broom)

# Import data
selection <- read_excel("databases/selection.xlsx", sheet = "data")

# Keep complete cases for the variables used in the model
selection_model <- selection %>%
  filter(!is.na(LeaverYr2),
         !is.na(Year1performanceRating))

# Logistic regression
# HR is omitted as the reference job function
model_p5 <- glm(
  LeaverYr2 ~
    Gender + EducationHighest + UGYN + WorkExperience +
    ACPersonalityO + ACPersonalityC + ACPersonalityE + ACPersonalityA + ACPersonalityN +
    ACRatingINTCOMPA + ACRatingINTCOMPB + ACRatingINTCOMPC +
    ACRatingINTCOMPD + ACRatingINTCOMPE +
    ACRatingAPTnumerical + ACRatingAPTverbal +
    InductionDay + InductionWeek + OnBoardingBuddy +
    FinanceDummyV + MarketingDummyV + SalesDummyV +
    RiskDummyV + LegalDummyV + OperationsDummyV +
    Year1performanceRating,
  data = selection_model,
  family = binomial(link = "logit")
)

summary(model_p5)

# Odds ratios with 95% CI
p5_results <- broom::tidy(model_p5, conf.int = TRUE, exponentiate = TRUE) %>%
  arrange(p.value)

p5_results

```
A binary logistic regression was conducted to predict LeaverYr2 using demographic variables, assessment-centre measures, aptitude scores, onboarding variables, job-function controls, and Year 1 performance rating. The model fit the data significantly better than the null model, as shown by the substantial reduction in deviance from 380.76 to 263.62, indicating that the predictors jointly provide meaningful information about early turnover. Based on the deviance reduction, the model explains approximately 31% of the variation in turnover risk, which represents a reasonably strong result for an HR attrition model.

The strongest predictor in the model was Attend Induction Day, which was associated with substantially lower odds of leaving by Year 2 (OR = 0.14, 95% CI [0.04, 0.41], p < .001). This indicates that employees who attended Induction Day had approximately 86% lower odds of early turnover, holding the other variables constant. Significant positive risk factors included Gender (OR = 4.37, p = .013), Assessment Centre Personality Neuroticism (OR = 1.02, p = .033), and several job-function controls, particularly Finance, Marketing, Risk, and Operations, all of which showed substantially higher odds of turnover relative to the omitted reference function. However, these job-function effects should be interpreted cautiously because their confidence intervals are wide, indicating uncertainty around the precise magnitude of the effect.

Importantly, Year 1 performance rating was not a significant predictor of leaving by Year 2 (OR = 1.09, p = .721). Most assessment-centre competency variables were also non-significant, as were UG status, education level, work experience, verbal aptitude, Induction Week, and OnBoardingBuddy. Overall, these findings suggest that early turnover in this sample is not primarily explained by lower first-year performance, but instead appears more closely related to onboarding participation, personality-related vulnerability, and structural differences across job functions.

### Recommendations for reducing graduate attrition
The most immediate recommendation is to strengthen and monitor Induction Day participation, since it was the clearest protective factor in the model. If attendance is inconsistent, optional, or unevenly implemented across functions, the organization should review this urgently. The size of the observed association suggests that early integration may play an important role in whether graduates remain with the firm.

Second, the company should investigate why turnover risk is markedly higher in Finance, Marketing, Risk, and Operations than in the reference function. Since Year 1 performance rating was not significant, these differences do not appear to be explained simply by weaker employees leaving. HR should therefore examine role demands, manager support, workload, expectations, and early career experience within those business areas.

Third, the significant association between Gender and attrition suggests the need for a closer review of whether some groups experience the graduate journey differently. This should not be treated as proof of causality, but it is strong enough to justify follow-up analysis of support, inclusion, progression opportunities, and work climate. Finally, because Neuroticism was positively associated with leaving risk, the organization may benefit from strengthening resilience-building, check-ins, and tailored support during the early employment period. Overall, the most effective retention strategy is likely to focus on onboarding quality and function-specific experience, rather than assuming that early leavers are simply poor performers.
