---
title: "Recruitment and Selection Analytics" 
subtitle: "Groupwork Use Case 5 - People Analytics"
author: "Substitute Group_X by your Group number in the name of the .Rmd file"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
editor_options: 
  markdown: 
    wrap: 72
---

# Intro

In this final workshop, you will apply analytics techniques to
recruitment and selection data to identify potential biases, evaluate
the fairness of selection decisions, and assess whether assessment
centre data can predict performance and turnover. You will use
chi-square tests, logistic and linear regression models to uncover
hidden patterns and build evidence-based recommendations.

# Setup

Within the project directory, create the following folders if they do
not already exist:

-   **Databases**: to store all data sources provided in Moodle (Excel
    files).\
-   **data**: to store any ad-hoc data created during the analysis. The
    native R data format is .rds.\
-   **report**: to store the HTML output file generated when knitting
    the R Markdown file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
```

# Prompt 1: Exploring Diversity in the Applicant Pool

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Recode gender and UG variables.\
-   Generate frequency tables for Gender, UG, Shortlisted, Interviewed,
    Offered, Accepted, Joined.\
-   Interpret patterns in diversity at each recruitment stage.\

### Deliverables

-   Table of frequencies.\
-   Initial diagnosis: Where in the pipeline are UG or female applicants
    being lost?

## Solution

### Table of frequencies

```{r prompt1.1}
# Load libraries
library(readxl)
library(dplyr)
library(janitor)

# Import dataset
df1 <- read_excel("Databases/recruitment_applicants_2023.xlsx")

# Clean names
df1 <- clean_names(df1)

# Check names
names(df1)
names(df1)
glimpse(df1)
# Recode variables
# Recode variables
df1 <- df1 %>%
  mutate(
    gender_rec = case_when(
      gender %in% c(1, "1", "F", "Female") ~ "Female",
      gender %in% c(2, "2", "M", "Male") ~ "Male",
      TRUE ~ "Other/Unknown"
    ),
    ug_rec = case_when(
      u_gyn %in% c(1, "1", "Y", "Yes") ~ "UG",
      u_gyn %in% c(2, "2", "N", "No") ~ "Non-UG",
      TRUE ~ "Unknown"
    )
  )

# Frequency tables
freq_gender <- tabyl(df1$gender_rec)
freq_ug <- tabyl(df1$ug_rec)
freq_shortlisted <- tabyl(df1$shortlisted_ny)
freq_interviewed <- tabyl(df1$interviewed)
freq_offered <- tabyl(df1$offer_ny)
freq_accepted <- tabyl(df1$accept_ny)
freq_joined <- tabyl(df1$join_yn)

freq_gender
freq_ug
freq_shortlisted
freq_interviewed
freq_offered
freq_accepted
freq_joined

# Print results

```

The results provide a clear overview of how the recruitment process
operates and where the main issues may lie.

First, the applicant pool is already imbalanced in terms of gender, with
only around 28% female candidates compared to over 70% male candidates.
This suggests that part of the diversity challenge originates before the
selection process itself, likely at the attraction or application stage.
In contrast, the proportion of candidates from underrepresented groups
is more balanced, which makes it possible to meaningfully assess whether
the selection process treats these groups fairly.

Looking at the progression through the recruitment stages, the data
shows a strong filtering effect in the early phases. Only about 31% of
applicants are shortlisted, and less than 20% reach the interview stage.
This indicates that the majority of candidates are eliminated very early
in the process. However, once candidates reach the interview stage, the
process becomes more selective but also more efficient, with
approximately half of interviewed candidates receiving an offer.
Finally, around 6% of the original applicant pool ends up accepting and
joining the organization.

This pattern suggests that the main bottleneck occurs during the
shortlisting and interview selection phases. These stages are typically
more subjective and involve greater human judgment, which increases the
risk of inconsistency and potential bias. According to recruitment
analytics theory, such stages are particularly vulnerable to effects
like first impressions or unconscious stereotyping, which may influence
decisions even when formal criteria are in place .

Another important aspect is the structure of the data itself. The large
number of missing values in later stages (offers, acceptance, and
joining) is expected, as only candidates who progress through earlier
stages are observed in subsequent ones. However, this means that any
deeper analysis must account for this conditional structure, otherwise
the conclusions could be misleading.

Overall, the results indicate that the recruitment process is highly
selective at the beginning and more consistent in later stages. While
this may reflect an efficient filtering mechanism, it also raises
concerns about whether early decisions are being made fairly and
systematically. Further statistical analysis, such as chi-square tests
or regression models, is necessary to determine whether these patterns
reflect true merit-based selection or potential bias against certain
groups. \### Initial diagnosis: Where in the pipeline are UG or female
applicants being lost? The data indicates that the largest drop in
candidates occurs during the transition from application to
shortlisting, and then again before the interview stage. Since these are
the most restrictive phases of the process, they are also the most
likely points where female or UG applicants could be disproportionately
filtered out.

Although the current results do not yet provide a breakdown by gender or
UG at each stage, the structure of the pipeline strongly suggests that
any loss of diversity is happening early rather than later. Once
candidates reach the interview stage, progression rates improve
significantly, making it less likely that major disparities occur at
that point.

Therefore, the initial diagnosis is that shortlisting is the critical
stage where potential bias or unequal outcomes may arise. Further
analysis, such as comparing selection rates by gender and UG, would be
necessary to confirm whether these groups are being systematically
disadvantaged

# Prompt 2: Testing for Bias in Shortlisting and Offers

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Use cross-tabulation and chi-square tests to compare: Gender ×
    Shortlisted; UG × Shortlisted; Gender × Offer; UG × Offer.\
-   Identify significant imbalances.\

### Deliverables

-   Chi-square test results with interpretation.\
-   Summary: Are the recruitment decisions fair and unbiased?\

## Solution

### Chi-square test results with interpretation

```{r prompt2.1}
library(readxl)
df2 <- read_excel("Databases/recruitment_applicants_2023.xlsx")
```

```{r}
names(df2)
```

```{r}
str(df2)
```

```{r prompt2.2}
# Quick checks
summary(df2)
```

```{r prompt2.3}
# 2. Recodificar las variables relevantes para que las tablas sean legibles
df2 <- df2 %>%
  mutate(
    Gender = factor(Gender, levels = c(1, 2), labels = c("Male", "Female")),
    UGyn = factor(UGyn, levels = c(1, 2), labels = c("Yes", "No")),
    ShortlistedNY = factor(ShortlistedNY, levels = c(0, 1), labels = c("Rejected", "Shortlisted")),
    OfferNY = factor(OfferNY, levels = c(0, 1), labels = c("No Offer", "Made Offer"))
  )
```

```{r prompt2.4}
# 3. Generar Tablas Cruzadas y Pruebas Chi-cuadrado

# Gender × Shortlisted
cat("\n--- Gender x Shortlisted ---\n")
tab_gen_short <- table(df2$Gender, df2$ShortlistedNY)
print(tab_gen_short)
print(chisq.test(tab_gen_short))
```

```{r prompt2.5}
# UG × Shortlisted
cat("\n--- UG (Underrepresented Group) x Shortlisted ---\n")
tab_ug_short <- table(df2$UGyn, df2$ShortlistedNY)
print(tab_ug_short)
print(chisq.test(tab_ug_short))
```

```{r prompt2.6}
# Gender × Offer
cat("\n--- Gender x Offer ---\n")
tab_gen_offer <- table(df2$Gender, df2$OfferNY)
print(tab_gen_offer)
print(chisq.test(tab_gen_offer))
```

```{r prompt2.7}
# UG × Offer
cat("\n--- UG (Underrepresented Group) x Offer ---\n")
tab_ug_offer <- table(df2$UGyn, df2$OfferNY)
print(tab_ug_offer)
print(chisq.test(tab_ug_offer))
```

### Summary: Are the recruitment decisions fair and unbiased?

# Prompt 3: Predicting Shortlisting Using Demographics

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all
    fields are included in the 'variable key' worksheet within the Excel
    file).\
-   Run logistic regression with Gender and UG as predictors and
    ShortlistedNY as target.\
-   Interpret odds ratios and model fit (Nagelkerke R²).\
-   Comment on potential compound bias (e.g., female & UG status).

### Deliverables

-   Model output.\
-   Odds ratios and statistical significance.\
-   Recommendations to improve fairness in shortlisting.\

## Solution

### Model output

```{r prompt3.1}


```

### Odds ratios and statistical significance

```{r prompt3.2}


```

### Recommendations to improve fairness in shortlisting

# Prompt 4: Validating Selection Tools Using Performance Ratings

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are
    included in the 'variable key' worksheet within the Excel file).\
-   Predict Year 1 performance rating using demographic, personality,
    competency, aptitude, and onboarding variables.\
-   Use linear regression with standardized coefficients.\
-   Interpret model fit and variable importance.

### Deliverables

-   Regression summary with top predictors.\
-   Evaluation of selection validity: Which tools truly predict
    performance?\

## Solution

### Regression summary with top predictors

```{r prompt4.1}
library(readxl)
library(dplyr)
library(lm.beta)
library(broom)
library(car)

# Import data
selection <- read_excel("Databases/selection.xlsx", sheet = "data")

# Keep only observations with Year 1 performance available
selection_model <- selection %>%
  filter(!is.na(Year1performanceRating))

# Linear regression
# Omit HRDummyV so HR becomes the reference category
model_p4 <- lm(
  Year1performanceRating ~
    Gender + EducationHighest + UGYN + WorkExperience +
    ACPersonalityO + ACPersonalityC + ACPersonalityE + ACPersonalityA + ACPersonalityN +
    ACRatingINTCOMPA + ACRatingINTCOMPB + ACRatingINTCOMPC +
    ACRatingINTCOMPD + ACRatingINTCOMPE +
    ACRatingAPTnumerical + ACRatingAPTverbal +
    InductionDay + InductionWeek + OnBoardingBuddy +
    FinanceDummyV + MarketingDummyV + SalesDummyV +
    RiskDummyV + LegalDummyV + OperationsDummyV,
  data = selection_model
)

# Standardized coefficients
model_p4_beta <- lm.beta(model_p4)

# Full summary
summary(model_p4_beta)

# Tidy table with standardized coefficients
p4_results <- broom::tidy(model_p4_beta) %>%
  mutate(std_beta = c(NA, lm.beta(model_p4)$standardized.coefficients[-1])) %>%
  arrange(desc(abs(std_beta)))

p4_results

# Optional: multicollinearity check
car::vif(model_p4)

```

**Leyend:** *ACRatingINTCOMPA = Assessment Centre Competency, Technical
ACRatingINTCOMPB = Assessment Centre Competency, Team Player
ACRatingINTCOMPC = Assessment Centre Competency Critical Thought
ACRatingINTCOMPD = Assessment Centre Competency Business Aware,
ACRatingINTCOMPE = Assessment Centre Competency Drive Innovation*

A multiple linear regression was conducted to predict **Year 1
performance rating** from demographic variables, personality measures,
assessment-centre competency ratings, aptitude test scores, onboarding
variables, and job-function controls. The overall model was
statistically significant, **F(25, 292) = 23.08, p \< .001**, indicating
that the predictors jointly explained a meaningful proportion of
variance in performance outcomes. The model showed strong explanatory
power, with **R² = 0.664** and **adjusted R² = 0.635**, meaning that
around 64% of the variance in Year 1 performance ratings was accounted
for by the included variables.

Examining the standardized coefficients, the strongest significant
predictors in the full model were **Gender** (β = -0.362, p \< .001),
**Legal job function** (β = 0.357, p \< .001), **Marketing job
function** (β = 0.224, p = .005), **Risk job function** (β = 0.219, p \<
.001), **Assessment Centre Personality Extraversion** (β = 0.179, p =
.006), **Assessment Centre Competency Technical** (β = 0.155, p = .009),
**Assessment Centre Competency Business Aware** (β = 0.150, p = .002),
**Assessment Centre Personality Conscientious** (β = 0.144, p = .003),
**Assessment Centre Competency Drive Innovation** (β = 0.123, p = .038),
and **Attend Induction Day** (β = 0.102, p = .012). In contrast, **UG
status**, **work experience**, **Openness**, **Agreeableness**,
**Neuroticism**, **Team Player**, **Critical Thought**, **Numerical
Aptitude Test Score**, **Verbal Reasoning Aptitude Test Score**,
**Induction Week**, and **OnBoardingBuddy** were not statistically
significant predictors in the full model.

### Evaluation of selection validity: Which tools truly predict performance?

To answer the validity question properly, it is important to distinguish
between **contextual control variables** and actual **selection tools**.
Although **Gender** and several **job-function dummies** were among the
strongest predictors in the model, these variables do not represent
selection methods themselves. Instead, they indicate that performance
ratings vary across demographic coding and business areas, so they
should be interpreted as contextual influences rather than evidence of
selection validity.

Focusing specifically on the selection-related measures, the clearest
evidence of predictive validity comes from selected **personality** and
**competency** indicators. In particular, **Extraversion** and
**Conscientiousness** were both positive and significant predictors of
Year 1 performance, suggesting that these traits are meaningfully
associated with stronger early job outcomes. In addition, the competency
ratings for **Technical**, **Business Aware**, and **Drive Innovation**
all made significant positive contributions, indicating that these
assessment-centre dimensions successfully capture capabilities linked to
later performance. Among onboarding measures, only **Attend Induction
Day** showed a significant effect, while **Induction Week** and
**OnBoardingBuddy** did not.

Overall, the findings suggest that the organization’s selection system
is **partially valid rather than uniformly valid**. Some tools clearly
contribute to predicting future performance, especially selected
personality dimensions and specific competency ratings. However, other
measures appear to add little incremental value once the full model is
taken into account. Most notably, the **numerical** and **verbal
aptitude tests** were not significant predictors in this analysis, nor
were the competency dimensions **Team Player** and **Critical Thought**.
This implies that the organization should place greater emphasis on the
measures that demonstrate clear predictive value, while critically
reviewing whether weaker tools meaningfully improve hiring decisions. In
short, the results support the validity of parts of the assessment
process, but they do not justify treating the entire assessment battery
as equally effective.

# Prompt 5: Predicting Early Turnover

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are
    included in the 'variable key' worksheet within the Excel file).\
-   Run logistic regression to predict LeaverYr2.\
-   Use demographics, assessments, onboarding, and performance rating.\
-   Evaluate odds ratios and risk factors.

### Deliverables

-   Model output and interpretation.\
-   Recommendations for reducing graduate attrition.\

## Solution

### Model output and interpretation

```{r prompt5.1}
library(readxl)
library(dplyr)
library(broom)

# Import data
selection <- read_excel("databases/selection.xlsx", sheet = "data")

# Keep complete cases for the variables used in the model
selection_model <- selection %>%
  filter(!is.na(LeaverYr2),
         !is.na(Year1performanceRating))

# Logistic regression
# HR is omitted as the reference job function
model_p5 <- glm(
  LeaverYr2 ~
    Gender + EducationHighest + UGYN + WorkExperience +
    ACPersonalityO + ACPersonalityC + ACPersonalityE + ACPersonalityA + ACPersonalityN +
    ACRatingINTCOMPA + ACRatingINTCOMPB + ACRatingINTCOMPC +
    ACRatingINTCOMPD + ACRatingINTCOMPE +
    ACRatingAPTnumerical + ACRatingAPTverbal +
    InductionDay + InductionWeek + OnBoardingBuddy +
    FinanceDummyV + MarketingDummyV + SalesDummyV +
    RiskDummyV + LegalDummyV + OperationsDummyV +
    Year1performanceRating,
  data = selection_model,
  family = binomial(link = "logit")
)

summary(model_p5)

# Odds ratios with 95% CI
p5_results <- broom::tidy(model_p5, conf.int = TRUE, exponentiate = TRUE) %>%
  arrange(p.value)

p5_results

```
A binary logistic regression was conducted to predict LeaverYr2 using demographic variables, assessment-centre measures, aptitude scores, onboarding variables, job-function controls, and Year 1 performance rating. The model fit the data significantly better than the null model, as shown by the substantial reduction in deviance from 380.76 to 263.62, indicating that the predictors jointly provide meaningful information about early turnover. Based on the deviance reduction, the model explains approximately 31% of the variation in turnover risk, which represents a reasonably strong result for an HR attrition model.

The strongest predictor in the model was Attend Induction Day, which was associated with substantially lower odds of leaving by Year 2 (OR = 0.14, 95% CI [0.04, 0.41], p < .001). This indicates that employees who attended Induction Day had approximately 86% lower odds of early turnover, holding the other variables constant. Significant positive risk factors included Gender (OR = 4.37, p = .013), Assessment Centre Personality Neuroticism (OR = 1.02, p = .033), and several job-function controls, particularly Finance, Marketing, Risk, and Operations, all of which showed substantially higher odds of turnover relative to the omitted reference function. However, these job-function effects should be interpreted cautiously because their confidence intervals are wide, indicating uncertainty around the precise magnitude of the effect.

Importantly, Year 1 performance rating was not a significant predictor of leaving by Year 2 (OR = 1.09, p = .721). Most assessment-centre competency variables were also non-significant, as were UG status, education level, work experience, verbal aptitude, Induction Week, and OnBoardingBuddy. Overall, these findings suggest that early turnover in this sample is not primarily explained by lower first-year performance, but instead appears more closely related to onboarding participation, personality-related vulnerability, and structural differences across job functions.

### Recommendations for reducing graduate attrition
The most immediate recommendation is to strengthen and monitor Induction Day participation, since it was the clearest protective factor in the model. If attendance is inconsistent, optional, or unevenly implemented across functions, the organization should review this urgently. The size of the observed association suggests that early integration may play an important role in whether graduates remain with the firm.

Second, the company should investigate why turnover risk is markedly higher in Finance, Marketing, Risk, and Operations than in the reference function. Since Year 1 performance rating was not significant, these differences do not appear to be explained simply by weaker employees leaving. HR should therefore examine role demands, manager support, workload, expectations, and early career experience within those business areas.

Third, the significant association between Gender and attrition suggests the need for a closer review of whether some groups experience the graduate journey differently. This should not be treated as proof of causality, but it is strong enough to justify follow-up analysis of support, inclusion, progression opportunities, and work climate. Finally, because Neuroticism was positively associated with leaving risk, the organization may benefit from strengthening resilience-building, check-ins, and tailored support during the early employment period. Overall, the most effective retention strategy is likely to focus on onboarding quality and function-specific experience, rather than assuming that early leavers are simply poor performers.
