---
# this is the Rmarkdown header. Use the YAML syntax. ---YAML block ---.
title: "Recruitment and Selection Analytics" 
subtitle: "Groupwork Use Case 5 - People Analytics"
author: "Substitute Group_X by your Group number in the name of the .Rmd file"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide

# the following lines tell R to store the output in the subdirectory reports
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(dirname(inputFile),"report" )) }) 
---

# Intro

In this final workshop, you will apply analytics techniques to recruitment and selection data to identify potential biases, evaluate the fairness of selection decisions, and assess whether assessment centre data can predict performance and turnover. You will use chi-square tests, logistic and linear regression models to uncover hidden patterns and build evidence-based recommendations.

# Setup

Within the project directory, create the following folders if they do not already exist:

- **Databases**: to store all data sources provided in Moodle (Excel files).\
- **data**: to store any ad-hoc data created during the analysis. The native R data format is .rds.\
- **report**: to store the HTML output file generated when knitting the R Markdown file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
```

# Prompt 1: Exploring Diversity in the Applicant Pool

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Recode gender and UG variables.\
-   Generate frequency tables for Gender, UG, Shortlisted, Interviewed, Offered, Accepted, Joined.\
-   Interpret patterns in diversity at each recruitment stage.\

### Deliverables

-   Table of frequencies.\
-   Initial diagnosis: Where in the pipeline are UG or female applicants being lost?

## Solution

### Table of frequencies

```{r prompt1.1}
# Load libraries
library(readxl)
library(dplyr)
library(janitor)

# Import dataset
data <- read_excel("Databases/recruitment_applicants_2023.xlsx")

# Clean names
data <- clean_names(data)

# Check names
names(data)
names(data)
glimpse(data)
# Recode variables
# Recode variables
data <- data %>%
  mutate(
    gender_rec = case_when(
      gender %in% c(1, "1", "F", "Female") ~ "Female",
      gender %in% c(2, "2", "M", "Male") ~ "Male",
      TRUE ~ "Other/Unknown"
    ),
    ug_rec = case_when(
      u_gyn %in% c(1, "1", "Y", "Yes") ~ "UG",
      u_gyn %in% c(2, "2", "N", "No") ~ "Non-UG",
      TRUE ~ "Unknown"
    )
  )

# Frequency tables
freq_gender <- tabyl(data$gender_rec)
freq_ug <- tabyl(data$ug_rec)
freq_shortlisted <- tabyl(data$shortlisted_ny)
freq_interviewed <- tabyl(data$interviewed)
freq_offered <- tabyl(data$offer_ny)
freq_accepted <- tabyl(data$accept_ny)
freq_joined <- tabyl(data$join_yn)

freq_gender
freq_ug
freq_shortlisted
freq_interviewed
freq_offered
freq_accepted
freq_joined

# Print results

```

### Initial diagnosis: Where in the pipeline are UG or female applicants being lost?


# Prompt 2: Testing for Bias in Shortlisting and Offers

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Use cross-tabulation and chi-square tests to compare: Gender × Shortlisted; UG × Shortlisted; Gender × Offer; UG × Offer.\
-   Identify significant imbalances.\


### Deliverables

-   Chi-square test results with interpretation.\
-   Summary: Are the recruitment decisions fair and unbiased?\

## Solution

### Chi-square test results with interpretation

```{r prompt2.1}


```

### Summary: Are the recruitment decisions fair and unbiased?


# Prompt 3: Predicting Shortlisting Using Demographics

### Tasks

-   Import 'recruitment_applicants_2023.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Run logistic regression with Gender and UG as predictors and ShortlistedNY as target.\
-   Interpret odds ratios and model fit (Nagelkerke R²).\
-   Comment on potential compound bias (e.g., female & UG status).

### Deliverables

-   Model output.\
-   Odds ratios and statistical significance.\
-   Recommendations to improve fairness in shortlisting.\

## Solution

### Model output

```{r prompt3.1}


```

### Odds ratios and statistical significance

```{r prompt3.2}


```

### Recommendations to improve fairness in shortlisting


# Prompt 4: Validating Selection Tools Using Performance Ratings

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Predict Year 1 performance rating using demographic, personality, competency, aptitude, and onboarding variables.\
-   Use linear regression with standardized coefficients.\
-   Interpret model fit and variable importance.

### Deliverables

-   Regression summary with top predictors.\
-   Evaluation of selection validity: Which tools truly predict performance?\


## Solution

### Regression summary with top predictors

```{r prompt4.1}


```

### Evaluation of selection validity: Which tools truly predict performance?

# Prompt 5: Predicting Early Turnover

### Tasks

-   Import 'selection.xlsx' into R (definitions of all fields are included in the 'variable key' worksheet within the Excel file).\
-   Run logistic regression to predict LeaverYr2.\
-   Use demographics, assessments, onboarding, and performance rating.\
-   Evaluate odds ratios and risk factors.

### Deliverables

-   Model output and interpretation.\
-   Recommendations for reducing graduate attrition.\

## Solution

### Model output and interpretation

```{r prompt5.1}


```

### Recommendations for reducing graduate attrition